---
title: "PA1_template.Rmd"
output: html_document
---

##1.Loading and processing the data
 
```{r Loadpackages}

library(ggplot2)
library(dplyr)
```

```{r Loaddata}

Activity <- read.csv(file = "activity.csv",na.strings =NA)  
summary(Activity)
head(Activity)
names(Activity)
Activity$date <- as.Date(Activity$date,"%Y-%m-%d")
head(Activity$date)

#Dpylr seems to be not removing NA.RM effectively. Hence removing
#them here
NActivity <-Activity[!is.na(Activity$steps),]

```
##2. Histogram of the total number of steps take each day

```{r Histogram}

TActivity <- NActivity %>% group_by(date) %>% summarize(Totsteps = sum(steps))

head(TActivity)
hist(TActivity$Totsteps,col ="Brown",main="Total steps taken each day",xlab="Total steps")

```

##3. Mean and median for number of steps taken

``` {r Meanmedian}
mean_act <- round(mean(TActivity$Totsteps),2)
mean_act
median_act <- median(TActivity$Totsteps)
median_act
#hist(TActivity$Totsteps,main="Total steps taken each day Mean (red) and Median (blue)",xlab="Total steps")
#abline(v=mean_act,col="red",lwd=1)
#abline(v=median_act,col="blue",lwd=1)

```
### The mean is `r mean_act` and median is `r median_act`

##4. Time series plot of the average number of steps 

```{r Timeseries}

TIActivity <- NActivity %>% group_by(interval) %>% summarise(avgsteps = mean(steps,na.rm=TRUE))

plot(TIActivity$interval,TIActivity$avgsteps,col="purple",type="l",xlab="Timeline",ylab="Average Steps",main="Time plot of average steps per day")

```

##5.Time interval

``` {r Timeinterval}


#Highest average interval and the average
avgti <- arrange(TIActivity,desc(avgsteps))[1,1]
avgst <- round(arrange(TIActivity,desc(avgsteps))[1,2],1)
avgti
avgst

```
### The highest average interval is `r avgti` and average steps
### are `r avgst` 

##6. Strategy for imputing missing values.
 - How many NA's are there?
 - Use mean of the time interval from above code.

``` {r Imputing}

#Total number of missing values.

sum(is.na(Activity$steps))


# Join the mean of the time interval
NewActivity <- left_join(Activity,TIActivity)

#Split it into na steps and !na. And replace avgsteps into steps
NewActivityna <- NewActivity[is.na(NewActivity$steps),]
NewActivitynna <- NewActivity[!is.na(NewActivity$steps),]

NewActivityna$steps <-NewActivityna$avgsteps

ActivityNew <- rbind(NewActivityna,NewActivitynna) 


sum(is.na(ActivityNew$steps))

```

##7. Histogram without missing values

``` {r newhist}
TActivityNew <- ActivityNew %>% group_by(date) %>% summarize(Totsteps = sum(steps,na.rm=TRUE))

hist(TActivityNew$Totsteps,col = "blue",main="Total steps taken each day with imputed values",xlab="Total steps")

nmean_act <- round(mean(TActivityNew$Totsteps),0.01)
nmean_act
nmedian_act <- median(TActivityNew$Totsteps)
nmedian_act

```

### OLD mean is `r mean_act` and median is `r median_act`
### NEW mean is `r nmean_act` and median is `r nmedian_act`

##8) 5 min interval plots  across weekends and weekdays

``` {r week}

cweekd <- function (x){
  
  if (x=="Sunday"|x=="Saturday"){
      return ("Weekend")
        }
  else {
    return ("Weekday")
      }
} 

ActivityNew$weekd <- weekdays(ActivityNew$date)
ActivityNew$weekd <- sapply(ActivityNew$weekd,cweekd)

TIActivitynew <- ActivityNew %>% group_by(interval,weekd) %>% summarise(avgstepsnew = mean(steps,na.rm=TRUE))


ggplot(TIActivitynew,aes(interval,avgstepsnew,color=weekd))+geom_line() +xlab("Time Interval") +
  ylab(expression("Average steps")) +
  ggtitle("Average steps per time interval by weekday type")



#plot ()

```